---
title: "Readme"
format: gfm
execute: 
  echo: true
  message: false
  warning: false
---

## Purpose

Collect soil quality data and aggregate up to NUTS2 level.

## Download data

You can download the data in excel format at [this link]()

## Context

Kerstin asked if I could get the soil quality data to supplement the dataset in her course.

### Data

Data on wheat suitability from FAO-GAEZ 4, available at the [following link](https://gaez-data-portal-hqfao.hub.arcgis.com/pages/data-viewer).

```{r}
library(tidyverse)
theme_set(theme_light())
library(gt)

tribble(
  ~Field, ~Value,
  "Theme", "Suitability and Attainable Yield",
  "Sub-theme", "Suitability Class",
  "Variable name", "Crop suitability index in classes",
  "Time period", "1981-2010",
  "Climate data source", "CRUTS32",
  "Crop", "Wheat",
  "Water supply", "Rainfed",
  "Input level", "Low"
) %>% 
  gt::gt() %>% 
  tab_header(title = "FAO-GAEZ 4 metadata") %>% 
  as_raw_html(inline_css = FALSE)
```

It is in a raster format - where the earth is broken into tiles.

This is what it looks like:

```{r}
library(stars)

tif <- read_stars(here::here("data", "siLr_whe_class.tif"))

ggplot() +
  geom_stars(data = tif)
```

We want to aggregate this to the nuts2 levels used in the roses wolf database.

### NUTS2 shapefile

The NUTS2 shapefiles are from [Eurostat](https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units/nuts)

We're going to work in WGS84 for convenience (EPSG: 4326).

We now need to aggregate up the NUTS2 regions which are joined together in the Roses Wolf database.

They are:

```{r}
tribble(
  ~NUTS2_regions, ~Aggregation,
  "AT12+AT13", "AT123",
  "DE71+DE72", "DE712",
  "DE91+DE92", "DE912"
) %>%
  gt() %>%
  tab_header(title = "Aggregations of NUTS2 regions")
```


```{r}
library(sf)

# map <- read_sf(here::here("maps", "regions_nuts2.shp")) %>% janitor::clean_names()

# map_tbl <- map %>% as_tibble()

# map_tbl %>%
#   ggplot(aes(geometry = geometry)) +
#   geom_sf() +
#   coord_sf()

## Making the map files smaller
# library(rmapshaper)
# map_simple <- ms_simplify(map, keep = 0.1,
#                                 keep_shapes = FALSE)
# 
# map_tbl_simple <- map_simple %>% as_tibble()
# 
# map_tbl_simple %>% write_rds(here::here("maps", "NUTS2_simple.rds"))

map_simple <- read_rds(here::here("maps", "NUTS2_simple.rds")) %>% 
  st_as_sf()

map_simple <- map_simple %>% 
  st_transform(crs = 4326)
```

We want to crop to only Europe:

```{r}
tif_bbox <- st_crop(tif, map_simple %>% st_bbox())

ggplot() +
  geom_stars(data = tif_bbox)
```

Now we want to calculate the averages within each polygon, along with the min and max values within each polygon.

```{r}
# devtools::install_github("michaeldorman/geobgu")
library(geobgu)

map_simple_avgs <-
  map_simple %>% mutate(
    suitability_class_mean = raster_extract(tif_bbox, map_simple, fun = mean, na.rm = TRUE),
    suitability_class_max = raster_extract(tif_bbox, map_simple, fun = max, na.rm = TRUE),
    suitability_class_min = raster_extract(tif_bbox, map_simple, fun = min, na.rm = TRUE)
  )
```

This is what the averaging procedure produces:

```{r}
map_simple_avgs %>% ggplot() +
  geom_sf(aes(fill = suitability_class_mean)) +
  scale_fill_viridis_c() +
  coord_sf() +
  labs(fill = "Soil suitability\nfor rainfed wheat\n(mean of class,\nlower is better)")
```

The distribution of minimum and maximum values is quite erratic. The plot below arranges the NUTS2 regions from lowest mean of suitability class to largest. The ribbon shows the 

```{r}
map_simple_avgs %>%
  st_set_geometry(NULL) %>%
  mutate(nuts_code = fct_reorder(nuts_code, suitability_class_mean)) %>%
  ggplot(aes(
    x = nuts_code, y = suitability_class_mean,
    ymin = suitability_class_min,
    ymax = suitability_class_max,
    group = 1
  )) +
  geom_line() +
  geom_ribbon(alpha = .3) +
  labs(
    x = "NUTS code",
    y = "Soil suitability index"
  )
```

Here is an example of 10 random regions.

```{r}
map_simple_avgs %>%
  slice_sample(n = 10) %>% 
  st_set_geometry(NULL) %>%
  knitr::kable()
```

### Join to Roses Wolf database

  "AT12+AT13", "AT123",
  "DE71+DE72", "DE712",
  "DE91+DE92", "DE912"

```{r}
df <- haven::read_dta(here::here("data", "RosesWolf_Regional_Fahad.dta"))

df <- df %>%
  distinct(id, nuts, region, country)

map_simple_avgs <- map_simple_avgs %>%
  st_set_geometry(NULL) %>%
  mutate(nuts = case_when(
    nuts_code == "AT123" ~ "AT12+AT13",
    nuts_code == "DE712" ~ "DE71+DE72",
    nuts_code == "DE912" ~ "DE91+DE92",
    TRUE ~ nuts_code
  ))

df_out <- df %>%
  left_join(map_simple_avgs) %>%
  select(-nuts_code)

# df_out %>% 
#   write.csv(here::here("data", "soil_suitability.csv"))
```


